Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-08-03 Final_Manuscript.tex
Worked on discussion

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Objectification in Action: Self- and Other-Objectification in Same-gender Interactions},
            pdfauthor={Randi L. Garcia, Kat Kyuchukova, Asha Hinson, \& Lanee Jung},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Objectification in Action: Self- and Other-Objectification in
Same-gender Interactions}
    \author{Randi L. Garcia\textsuperscript{1, 2}, Kat
Kyuchukova\textsuperscript{1}, Asha Hinson\textsuperscript{1}, \& Lanee
Jung\textsuperscript{1}}
    \date{}

@@

There is evidence that the experience of state self-objectification in
mixed-gender contexts (stranger and romantic) and within the self in any
moment (trying on a swimsuit) has negative consequences, but what about
in the context of interaction with other women? Is the
self-objectification experience in mixed-gender interactions the same as
that experienced in same-gender interactions? The ample research
demonstrating that the male gaze has a particularly detrimental effect
would suggest no (Gay \& Castano, 2010; Gervais, Vescio, \& Allen, 2011;
Saguy, Quinn, F Dovidio, \& Pratto, 2010, Calogero (2004)). Women do
objectify other women (Harsey \& Zurbriggen, 2020; Loughnan et al.,
2015a; Puvia \& Vaes, 2013), but when they do it, does it lead to
self-objectification in the same way men's objectification of women does
(Garcia et al., 2016)? If so, does the self-objectification experienced
in these same-gender interactions have the same negative consequences
for authenticity in that interaction?

@@

The current study tested whether the model of interpersonal
objectification and state self-objectification (SSO) used in Garcia et
al. (2016) replicates in a same-gender dyadic sample of women only.
Although past reserach has found that women do objectify other women
(Harsey \& Zurbriggen, 2020; Loughnan et al., 2015a; Puvia \& Vaes,
2013), this is the first study to test if this other-objectification by
women is related to self-objectification in women. As hypothesized, the
current study did find a significant relationship bewteen a woman's
partner's report of having objectified her and her own post-interaction
feelings of self-objectification.

Where the results of the current study diverge most notably from the
results of studies testing interpersonal objectification among
mixed-gender dyads is the lack of evidence for relationships between SSO
and felt inauthenticity (Garcia et al., 2016). In addition, we found to
evidence of \emph{direct} or \emph{indirect} relationships between SSO
and cognitive functioning, relationship agency, and career aspirations.
Nor was there evidence of \emph{direct} or \emph{indirect} relationships
between other-objectification and any of these outcomes. This is
somewhat suprising given the extant research on interpersonal
other-ojectification and cognitive functioning {[}CITE that logel
study{]}, and the plentiful evidence linking SSO and cognitive
functioning {[}B. L. Fredrickson et al. (1998); CITE; CITE; CITE{]}.
This lack of evidence could potentially signal diverging processs
between women's experiences with interpersonal objectification from men
and interpersonal objectification from women. There is quite a bit of
evidence suggesting that the male gaze is particularly detrimental
{[}CITE; CITE; CITE{]}, and perhaps the self-objectification experienced
within an interaction with a women is not as detrimental as the
self-objectificatoin experienced within an interaction with a man. This
lack of evidence could also be due to the relatively small sample size.

What we did find, however,

As stated previously, we did not find a significant effect between actor
SSO and actor authenticity, which suggests that there is not sufficient
evidence to support the claim that partner objectification is the cause
for the diverse range of negative effects related to relationship
inauthenticity. However, we did observe a significant partner effect of
objectification on actor self-objectification, which does align with our
hypothesis that theorized objectification from other women can also
cause women to self-objectify just as they do within interactions
between male-female pairs.

\subsection{Sexism and Sexual
Discrimination}\label{sexism-and-sexual-discrimination}

@@

@@

@@

The results of this study demonstrate the complex and ambivalent nature
of female sexual objectification and additionally highlight the
psychological and social consequences of such objectification processes
on women's social relationships and well-being.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Objectification in Action: Self- and Other-Objectification in Same-gender Interactions},
            pdfauthor={Randi L. Garcia, Kat Kyuchukova, Asha Hinson, \& Laneé Jung},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Objectification in Action: Self- and Other-Objectification in
Same-gender Interactions}
    \author{Randi L. Garcia\textsuperscript{1, 2}, Kat
Kyuchukova\textsuperscript{1}, Asha Hinson\textsuperscript{1}, \& Laneé
Jung\textsuperscript{1}}
    \date{}

@@

There is evidence that the experience of state self-objectification in
mixed-gender contexts (stranger and romantic) and within the self in any
moment (trying on a swimsuit) has negative consequences, but what about
in the context of interaction with other women? Is the
self-objectification experience in mixed-gender interactions the same as
that experienced in same-gender interactions? The ample research
demonstrating that the male gaze has a particularly detrimental effect
would suggest no (Gay \& Castano, 2010; Gervais, Vescio, \& Allen, 2011;
Saguy, Quinn, F Dovidio, \& Pratto, 2010, Calogero (2004); Yilmaz \&
Bozo, 2019). Women do objectify other women (Harsey \& Zurbriggen, 2020;
Loughnan et al., 2015a; Puvia \& Vaes, 2013), but when they do it, does
it lead to self-objectification in the same way men's objectification of
women does (Garcia et al., 2016)? If so, does the self-objectification
experienced in these same-gender interactions have the same negative
consequences for authenticity in that interaction?

@@

The current study tested whether the model of interpersonal
objectification and state self-objectification (SSO) used in Garcia et
al. (2016) replicates in a same-gender dyadic sample of women only.
Although past reserach has found that women do objectify other women
(Harsey \& Zurbriggen, 2020; Loughnan et al., 2015a; Puvia \& Vaes,
2013), this is the first study to test if \emph{interpersonal}
other-objectification by women is related to state self-objectification
in their woman-identified interaction partners. As hypothesized, the
current study did find a significant relationship bewteen a woman's
partner's report of having objectified her and her own post-interaction
feelings of self-objectification. That is, there was a significant
partner effect of other-objectification on SSO. This effect extends the
equivalent relationship found in mixed-gender interactions to the
context of same-gender interactions between women. Thus, it could be
that it is not only the male gaze, and male other-objectification that
is related to women's state self-objectification, but being objectified
by another woman can also results in SSO, at least in the context of a
scenario where they know they are being evaluated as a potential dating
partner.

As is the case in all correlational studies, we cannot be sure in the
causal direction. It could be that women's SSO causes them to be
objectified by their interaction partner. This interpretation in
theoreticall interesting given that, in both the Garcia et al. (2016)
study and in the current study, it is \emph{women's} SSO this relates to
being objectified by one's partner. However, Objectification Theory
(Roberts \& Fredrickson, 1997), as well as past experimental studies
{[}Saguy et al. (2010); CITE{]}, suggest that the causal flow is from
other-objectification to SSO. Although, importantly, the current study
found evidence for the partner effect of other-objectification on SSO,
there was a lack of evidence for the connections between SSO and
downstream negative consequences.

Where the results of the current study diverge most notably from the
results of studies testing interpersonal objectification among
mixed-gender dyads is the lack of evidence for relationships between SSO
and felt inauthenticity. In addition, we found no evidence of
\emph{direct} or \emph{indirect} relationships between SSO and cognitive
functioning, relationship agency, and career aspirations. Nor was there
evidence of \emph{direct} or \emph{indirect} relationships between
other-objectification and any of these outcomes. This is somewhat
suprising given the plentiful evidence linking SSO and cognitive
functioning (B. L. Fredrickson et al., 1998; Quinn, Chaudoir, \& Kallen,
2011) and the small, but extant research on interpersonal
other-ojectification and cognitive functioning (Garcia et al., 2016;
Logel et al., 2009). This lack of evidence could potentially signal
diverging processs between women's experiences with interpersonal
objectification from men and interpersonal objectification from women.
There is quite a bit of evidence suggesting that the male gaze is
particularly detrimental (Calogero, 2004; Gervais, Holland, \& Dodd,
2013; Gervais et al., 2011; Roberts \& Fredrickson, 1997), and perhaps
the self-objectification experienced within an interaction with a women
is not as harmful as the self-objectificatoin experienced within an
interaction with a man. However, as a strong note of caution, we need to
be careful not to interpret a null result as evidence of no
relationship.

The lack of evidence for a relationship between SSO and interaction
authenticity is surprising and again, should not be interpreted as
evidence of no realtionship. It should be noted that the estimate of
this relationship was very small, but in the negative direction, as
anticipated. If it is the case that there is a smaller (i.e., weaker)
connection between women's feelings of SSO and inauthenticity in
interactions with other women then in interactions with men, models of
interpersonal objectification, like the SIMO (Gervais et al., 2020),
could be extended by including inauthenticity as a potential mediating
factor in decisions to continue objectifying interactions. Further,
gender of the objectifyer/interaction partner could also be added to the
SIMO model to help extend our understanding of under what circumstances
other-objectification and SSO have negative concequences for women.

Although we did not find a connection bewteen SSO and authenticity we
did find significant positive relationships between authenticity and
relationship agency and career aspirations. The relationship between
authenticity and cognitive functioning was also estimated as positive,
but it was not statistically signfiicant. Again, due to the lack of
connection between SSO and authenticity, we found no evidence of
\emph{indirect} relationships between SSO and these outcome variables.
This evidence of a relationship between authenticity and the relational
outcome variables (i.e., relationship agency and career aspirations)
provides evidence that corrborates past findings that felt authenticity
in interactions is important for healthy relationship functioning
{[}Garcia et al. (2016); PWQ paper I reviewed{]}. Just as authenticity
has been found to be important in intergroup interactions {[}CITE; CITE;
CITE{]}, we again find more evidence here that distuptions in feelings
of authenticity can negatively impact relationships beyond the current
partner. Although the current study did not find a connection between
authenticity and SSO, this seems theoretically to be a natural
connection, and more work needs to be done to discover when and how SSO
leads to inauthenticity in interpersonal objectification situations.

@@

The sample size was small and combined across two higher education
institutions.

@@

\subsubsection{Sexual Objectification and the Objectifyer's
Gender}\label{sexual-objectification-and-the-objectifyers-gender}

The current sample contained a mixture of heterosexual and
non-heterosexual women, but all participants were asked to think about
and evaluate their partner as a potential data parter. While we think
that hetereosexual women are indeed able to do this with other women as
their target---indeed, there is evidence that they might do this readily
(Puvia \& Vaes, 2013; Strelan \& Hargreaves, 2005)---they might be more
apt to activate social comparison processes (Festinger, 1954) than women
who are sexually attracted to other women (non-heterosexual women). This
differential psychological process between women with differing
sexualities might have served to dampen our ability to detect
relationships, adding variability.

Previous research has found that when compared to heterosexual women,
lesbian women report less concern with physical appearance (Siever,
1994; Strong, Williamson, Netemeyer, \& Geer, 2000), and less
self-objectification (Brownlow, 1998; Noffsinger-Frazier, 2004). However
none of these studies examined the relationship between
self-objectification and experiences of sexual objectification. Thus, it
is unclear whether lesbians indeed experience similar levels of cultural
sexual objectification but internalize them less than heterosexual women
do. Consistent with previous reseach Hill and Fischer (2008) determined
lesbians existed less physical appearance concerns compared to
heterosexual women, however there was no difference found that lesbian
women self-objectify less than heterosexual women and they did not find
that sexual orientation moderates the relationshiotn between sexual
objecitifcantion and self-objectification This contradicts older
theoretical literature that suggests that lesbians internalize cultural
sexual objectification less than do heterosexual women (Brown, 1987;
LUM, 1994; Pitman, 1999; Siever, 1994)

However, since both men and women are socialized in a culture that
sexually objectifies women, both men and women may come to internalize
this socialization and sexually objectify women. Indeed, recent research
has found that the more women self-objectify, the more they objectify
other women (Loughnan et al., 2015a; Puvia \& Vaes, 2013), although not
to the degree exhibited by men; that is, men were found to objectify
women significantly more than women objectify other women (Strelan \&
Hargreaves, 2005).

\subsubsection{Sexism and Sexual
Discrimination}\label{sexism-and-sexual-discrimination}

@@

The results of this study demonstrate the complex and ambivalent nature
of female sexual objectification and additionally highlight the
psychological and social consequences of such objectification processes
on women's social relationships and well-being. We did not find a
significant effect between actor SSO and felt authenticity in the
interaction, which suggests that there is not sufficient evidence to
support the claim that partner objectification is the cause for the
diverse range of negative effects related to interaction inauthenticity.

@@

\hypertarget{ref-brown1987lesbians}{}
Brown, L. S. (1987). Lesbians, weight, and eating: New analyses and
perspectives. \emph{Lesbian Psychologies: Explorations and Challenges},
294--309.

\hypertarget{ref-brownlow1998relationship}{}
Brownlow, B. S. (1998). \emph{The relationship between objectification,
body image disturbances, and disordered eating: Investigating race,
socioeconomic status, acculturation and self-objectification as
mediators.} (PhD thesis). ProQuest Information \& Learning.

@@

\hypertarget{ref-festinger1954theory}{}
Festinger, L. (1954). A theory of social comparison processes.
\emph{Human Relations}, \emph{7}(2), 117--140.

@@

\hypertarget{ref-gervais2013my}{}
Gervais, S. J., Holland, A. M., \& Dodd, M. D. (2013). My eyes are up
here: The nature of the objectifying gaze toward women. \emph{Sex
Roles}, \emph{69}(11-12), 557--570.

@@

\hypertarget{ref-logel2009interacting}{}
Logel, C., Walton, G. M., Spencer, S. J., Iserman, E. C., Hippel, W.
von, \& Bell, A. E. (2009). Interacting with sexist men triggers social
identity threat among female engineers. \emph{Journal of Personality and
Social Psychology}, \emph{96}(6), 1089.

@@

\hypertarget{ref-lum1994lesbians}{}
LUM, E. (1994). Lesbians and physical appearance which model applies?
\emph{Lesbian and Gay Psychology: Theory, Research, and Clinical
Applications}, \emph{1}, 84.

@@

\hypertarget{ref-noffsinger2004objectification}{}
Noffsinger-Frazier, N. A. (2004). \emph{Objectification theory and
disordered eating: The impact of feminist identification,
internalization of sociocultural standards of appearance, and sexual
orientation}. The University of Memphis.

@@

\hypertarget{ref-pitman1999body}{}
Pitman, G. E. (1999). Body image, compulsory heterosexuality, and
internalized homophobia. \emph{Journal of Lesbian Studies}, \emph{3}(4),
129--139.

@@

\hypertarget{ref-quinn2011performance}{}
Quinn, D. M., Chaudoir, S. R., \& Kallen, R. W. (2011). Performance and
flow: A review and integration of self-objectification research.

@@

\hypertarget{ref-siever1994sexual}{}
Siever, M. D. (1994). Sexual orientation and gender as factors in
socioculturally acquired vulnerability to body dissatisfaction and
eating disorders. \emph{Journal of Consulting and Clinical Psychology},
\emph{62}(2), 252.

@@

\hypertarget{ref-strong2000eating}{}
Strong, S. M., Williamson, D. A., Netemeyer, R. G., \& Geer, J. H.
(2000). Eating disorder symptoms and concerns about body differ as a
function of gender and sexual orientation. \emph{Journal of Social and
Clinical Psychology}, \emph{19}(2), 240--255.

@@

\hypertarget{ref-yilmaz2019whose}{}
Yilmaz, T., \& Bozo, Ö. (2019). Whose gaze is more objectifying? An
experimental study of college women's state self-objectification, body
shame, negative mood, and body dissatisfaction. \emph{Mediterranean
Journal of Clinical Psychology}, \emph{7}(2).
